<span style="color:red;">ANTs:</span><br /> A retrospective

### 2015-05-20 background/background.Rmd
ENH:  Adding mni tutorial stuff.

@@

__Image mapping & perception: 1877-8__
<BR><BR>
Francis Galton: _Can we see criminality in the face?_

<!--
![](./background/figures/FrancisGalton2.jpg)
-->

@@

Speaking of criminality...<BR>
_Can we say anything about the U.S. Congress?_

![](./background/figures/congressPersons.png)
__Maybe he should have used [ANTs](https://github.com/ntustison/CongressionalFaceTemplates)?__

@@

__Image mapping & biology: 1917__
<BR><BR>
D''Arcy Thompson: _Comparison of related forms_

@@

__Image mapping & biology: Current__
<BR><BR>
_Look where we are now!_
<BR><BR>
![](./background/figures/r16xr85.png)

@@

__ANTs family tree__

@@

__Initial scope__

@@

__Diffeomorphisms__:plausible physical modeling

<div align="center"><img src="./figures/sillyputty.png" frameborder="0"></img></div>

"differentiable map with differentiable inverse"

__Fine-grained and flexible maps__

@@

 __Diffeomorphisms__: image parameterization in a metric space

@@

__General purpose library for multivariate image registration, segmentation & statistical analysis tools__

* 170,000+ lines of C++, 6$+$ years of work, 15+ collaborators.

* Generic mathematical methods that are tunable for
application specific domains.

* <span style="color:red;">no free lunch</span>: <span style="color:black;">_An algorithm must use prior knowledge about a problem to do well on that problem._</span>

<!--
* Deep testing on multiple platforms ... osx, linux, windows.

* Several "wins" in public knock-abouts ( [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496), [Murphy 2011](http://www.ncbi.nlm.nih.gov/pubmed/21632295), [SATA 2012 and 2013](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837555/), [BRATS 2013](http://martinos.org/qtim/miccai2013/proc_brats_2013.pdf), others )
-->

## Still quite a bit to do

__On documentation__

![](figures/Chainsaw.jpg)

<!--
Notes:
* ANTs documentation is only slightly better than what is shown here.
-->

Developers can be blind to doc deficiencies while users are blind to what we provide!

![](figures/blindmen.jpg)

>>>>

## Image mapping and perception: 1877

Francis Galton: _Can we see criminality in the face?_

_What about syphilis, mental illness?_

<!--

@@

![](./background/figures/FrancisGalton2.jpg)
-->

@@

##  Speaking of criminality...

_Can we say anything about the U.S. Congress?_

![](./background/figures/congressPersons.png)

__Maybe he should have used [ANTs](https://github.com/ntustison/CongressionalFaceTemplates)?__

@@

##  Image mapping & biology:  1917

D'Arcy Thompson: _Comparison of related forms_

@@

## Image mapping & biology: Current
<BR><BR>
_Look where we are now!_
<BR><BR>
![](./background/figures/r16xr85.png)

@@

## ANTs family tree

@@

## Initial scope

@@

## Diffeomorphisms:  Occam's razor modeling

![](./figures/sillyputty.png)

_differentiable map with differentiable inverse_

## Diffeomoprhisms:  fine-grained and flexible maps

@@

## Diffeomorphisms: image parameterization in a metric space

@@

### 2015-05-20 competitions/competitions.Rmd
ENH:  Adding mni tutorial stuff.

##

<span style="color:green;">__Competitions__</span>

@@

##

<span style="color:green;">__Klein 2009__</span>

@@

* 14 algorithms
* publicly available data
* different labeling protocols
* developers tune their own algorithms

## <span style="color:green;">Results</span>

@@

@@

## <span style="color:green;">EMPIRE 2010</span>

@@

* Alignment of lung boundaries,
* major fissures,
* annotated landmark pairs, and
* topology of displacement field.

##

@@

##

<span style="color:green;">__Multi-Atlas 2012__</span>

@@

##

@@

##

<span style="color:green;">SATA 2013</span>

@@

##

<span style="color:green;">__BRATS 2013__</span>

@@

##

@@

##

<span style="color:green;">__STACOM 2014__</span>

![](./competitions/figures/image_gallery.gif)

>>>>

## Algorithmic fight club

![](./competitions/figures/fightclub.jpg)

## International competitions

@@

## Klein, NeuroImage 2009

@@

<!--
* 14 algorithms
* publicly available data
* different labeling protocols
* developers tune their own algorithms
-->

## Klein results

@@

## Take-away message

@@

## EMPIRE 2010

@@

* Alignment of lung boundaries,

* major fissures,

* annotated landmark pairs, and

* topology of displacement field.

## EMPIRE 2010 results

@@

## Multi-atlas 2012

@@

## Multi-atlas 2012 results

@@

## SATA 2013

Three very different problem domains

@@

## BRATS 2013

@@

## BRATS 2013 results

@@

## STACOM 2014

![](./competitions/figures/image_gallery.jpg)

### 2015-05-20 evaluation/antsvfreesurfer2.Rmd
ENH:  Adding mni tutorial stuff.

# ANTs brain morphology<br><br>[<small>"*Large-scale evaluation of ANTs and FreeSurfer cortical thickness measurements*"</small>](http://www.ncbi.nlm.nih.gov/pubmed/24879923)

## Cortical thickness studies:
```{r, results='asis', echo = FALSE}
conditions <- c( "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender", "untreated male-to-female transsexuality",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                 "Tetris-playing ability in female adolescents", "traumatic brain injury" )

@@

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
  + used other tools such as FMRIB's Automated Segmentation Tool (FAST)
  + `KellySlater.cxx`
  * scripts "adequate for Penn data"
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

@@

<div align="center"><img height="350" src="./evaluation/figures/components.png" frameborder="0"></img></div>

@@

<div align="center"><img height="350" src="./evaluation/figures/templates.png" frameborder="0"></img></div>

* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.
* Acquisition and anonymization (e.g. defacing) protocols are often different.

## Template building (cont.)

<div align="center"><img height="350" src="./evaluation/figures/templatePriors.png" frameborder="0"></img></div>

Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.

## Brain extraction comparison<br>ANTs vs. FreeSurfer

<div align="center"><img height="380" src="./evaluation/figures/brainExtraction.png" frameborder="0"></img></div>

Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)

@@

<div align="center"><img height="350" src="./evaluation/figures/brainSegmentation.png" frameborder="0"></img></div>

Randomly selected healthy individuals.  Atropos gets good performance across ages.

@@

<div align="center"><img height="250" src="./evaluation/figures/corticalThicknessEstimation.png" frameborder="0"></img></div>

In contrast to FreeSurfer which warps coupled surface meshes to segment the gray matter, ANTs diffeomorphically registers the white matter to the combined gray/white matters while simultaneously estimating thickness.

@@

* Use age and gender (demographic information) in training/prediction paradigm
* Public data sets (IXI, NKI, Oasis, MMRR)
* $>$ 1200 subjects, age 7 to over 90 years old

## Prediction from cortical thickness data:<br>gender and age

<div align="center"><img height="350" src="./evaluation/figures/evaluation.png" frameborder="0"></img></div>

@@

<div align="center"><img height="500" src="./evaluation/figures/antsvfreesurfer_importance.png" frameborder="0"></img></div>

@@

## Regional importance comparison (cont.)

<div align="center"><img height="550" src="./evaluation/figures/antsvfreesurfer_regionalPlots.png" frameborder="0"></img></div>

## All software components are open source and part of the ANTs repository.

# Cortical thickness extensions

>>>>

---
title: ANTs brain morphology
author: Nick (UVa) and Brian (Penn)
bibliography: REFERENCES.bib
logo:    figures/antsLogoBase.png
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  structurebold
    highlight: tango
    incremental: yes
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
    toc: yes
  pdf_document:
    toc: true
    highlight: zenburn
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

```{r global options, include=FALSE}
library(rmarkdown)
library(knitr)
figpath<-'figures/'
opts_chunk$set(fig.width=9, fig.height=6, fig.path=figpath,
               warning=FALSE, message=FALSE)
```

## Cortical thickness studies

```{r, results='asis', echo = FALSE}
conditions <- c( "Tetris-playing ability", "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                  "traumatic brain injury", "untreated male-to-female transsexuality" )

@@

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
    + used other tools such as FMRIB Automated Segmentation Tool (FAST)
    + `KellySlater.cxx`
* pipedream scripts "adequate for Penn data."
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

## Sample results

![](./evaluation/figures/components.png)

@@

@@

![](./evaluation/figures/templates.png)

<!--
* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.
* Acquisition and anonymization (e.g. defacing) protocols are often different.
-->

## Template priors

![](./evaluation/figures/templatePriors.png)

<!--
Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.
-->

## Brain extraction comparison

![](./evaluation/figures/brainExtraction.png)

<!--
Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)
-->

@@

![](./evaluation/figures/brainSegmentation.png)

<!--
Randomly selected healthy individuals.  ```Atropos``` gets good performance across ages.
-->

@@

![](./evaluation/figures/corticalThicknessEstimation.png)

<!--
In contrast to FreeSurfer which warps coupled surface meshes to segment the gray matter, *ANTs* diffeomorphically registers the white matter to the combined gray/white matters while simultaneously estimating thickness.
-->

@@

## Predict age and gender

## Open science prinicples

![](./evaluation/figures/ageDistribution.png)

<!--
* Public data sets (IXI, NKI, OASIS, MMRR)
* $>$ 1200 subjects, age 4 to over 94 years old
-->

## Prediction from cortical thickness data

![](./evaluation/figures/evaluation.png)

## Age prediction per site

![](./evaluation/figures/agePredictionPerSite.png)

@@

![](./evaluation/figures/antsvfreesurfer_Importance.png)

@@

## Regional measurements

![](./evaluation/figures/antsvfreesurfer_regionalPlots.png)

### 2015-05-20 papers/papers.Rmd
ENH:  Adding mni tutorial stuff.

##

@@

##

__Symmetric Normalization:__  (_SyN_)

$\int_{t=0}^{0.5} \left(\|\mathbf{v}_1(x,t)\|_L^2 + \|\mathbf{v}_2(x,t)\|_L^2\right)dt$

$+ \|I\left(\phi_1(x,0.5)\right) - J_i\left(\phi_2(x,0.5)\right)\|^2$

@@

##

__Beyond original SyN__

@@

##

```antsRegistration```
<BR><BR>

~~~~~~~~~~~~~~~~~~~~~~~
$ antsRegistration --help

@@

     --help
          Print the help menu. Will also print values used on the current command line
          call.
~~~~~~~~~~~~~~~~~~~~~~~

##

__Template building__

##

@@

##

@@

##

@@

##

@@

##

@@

##

```antsMultivariateTemplateConstruction2.sh```

<BR>

~~~~~~~~~~~~~~~~~~~~~~
$ antsMultivariateTemplateConstruction2.sh

@@

--------------------------------------------------------------------------------------
Script by Nick Tustison
--------------------------------------------------------------------------------------
Apple XGrid support by Craig Stark
--------------------------------------------------------------------------------------
~~~~~~~~~~~~~~~~~~~~~~

##

__Nonparametric nonuniform intensity normalization (N3)__

@@

##

__N3 adoption issues__

@@

##

<!--
__N3 $\rightarrow$ N4<sup>*</sup>__

<p align = "middle"><small><sup>*</sup><u>N</u>ick's N3</small></p>
-->

__N4__

@@

* smaller spline distances (useful for higher magnet strengths)

@@

* faster execution times

@@

##

```N4BiasFieldCorrection```

<BR>

~~~~~~~~~~~~~~~~~~~~~~~~~~~
$ N4BiasFieldCorrection --help

@@

     --help
          Print the help menu.
          <VALUES>: 1
~~~~~~~~~~~~~~~~~~~~~~~~~~~

##

@@

<p align = "right"><small>Brian, MICCAI workshop</small></p>

@@

##

__Atropos__

@@

##

__Algorithmic workflow__<br>(Atropos + N4 $\rightarrow$ `antsAtroposN4.sh`)

##

```Atropos```

<BR>

~~~~~~~~~~~~~~~~~~~~~~~~~~~
$ Atropos --help

@@

     --help
          Print the help menu.
          <VALUES>: 1
~~~~~~~~~~~~~~~~~~~~~~~~~~~

>>>>

## Donoho?

@@

## Symmetric Normalization (SyN)

$\int_{t=0}^{0.5} \left(\|\mathbf{v}_1(x,t)\|_L^2 + \|\mathbf{v}_2(x,t)\|_L^2\right)dt + \|I\left(\phi_1(x,0.5)\right) - J_i\left(\phi_2(x,0.5)\right)\|^2$

@@

## Beyond original SyN

\small

@@

## ```antsRegistration```
<BR><BR>

```bash
$ antsRegistration --help

@@

     --help
          Print the help menu. Will also print values used on the current command line
          call.
```

## Template building:  creating the average Joe

@@

## "Attractiveness" $\rightarrow$ mental processing?

@@

##  What about brains?

@@

## Templates facilitate computation

@@

## Gender discernibility?

@@

## ```antsMultivariateTemplateConstruction2.sh```

```bash
$ antsMultivariateTemplateConstruction2.sh

@@

--------------------------------------------------------------------------------------
Script by Nick Tustison
--------------------------------------------------------------------------------------
Apple XGrid support by Craig Stark
--------------------------------------------------------------------------------------
```

## Nonparametric nonuniform intensity normalization (N3)

@@

## N3 adoption issues

@@

## N4

@@

* small spline distances (useful for higher magnet strengths)

@@

* fast execution times

@@

* tested nightly within the ITK dashboard system

## ```N4BiasFieldCorrection```

```bash
$ N4BiasFieldCorrection --help

@@

     --help
          Print the help menu.
          <VALUES>: 1
```

##  Atropos -- cutting the threads of fate

@@

@@

## Atropos:  flexible code base

@@

## ```Atropos```

```bash
$ Atropos --help

@@

     --help
          Print the help menu.
          <VALUES>: 1
```

## Atropos + N4 $\rightarrow$ `antsAtroposN4.sh`

### 2015-05-20 retrospective/neuroimage_fmri.Rmd
ENH:  Adding mni tutorial stuff.

@@

@@

@@

<!--
Notes:
* Additional interesting bits of information
* Logos
* Why a fish skeleton?
* "FOSSIL vs EFF ESS ELL"
* What is the ice cream cone for and what does it have to do with British humour?
-->

<BR>
<BR>
<BR>
<span style="color:red;">_"Contingency.
<BR>
AFNI is one result of unexpected events.
<BR>
I met a woman and fell in love."_</span>
<BR>

@@

## <span style="color:red;">ANTs--_begotten in SyN_</red

>>>>

## ANTs evolution

@@

## FreeSurfer, Bruce Fischl

@@

<!--

## Other juicy bits

@@

Notes:
* Additional interesting bits of information
* Logos
* Why a fish skeleton?
* "FOSSIL vs EFF ESS ELL"
* What is the ice cream cone for and what does it have to do with British humour?
-->

## NeuroImaging research as literature

"Contingency.
AFNI is one result of unexpected events.
I met a woman and fell in love."

@@

## ANTs--_begotten in SyN_

![](./retrospective/figures/Welcometotherealm.jpg)

### 2015-05-20 retrospective/retrospective.Rmd
ENH:  Adding mni tutorial stuff.

---
title: <span style="color:red;">ANTs:</span><br />A retrospective
author: Brian (<span style="color:blue;">PENN</span>) and <br />Nick (<span style="color:orange;">UVA</span>)
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

>>>>

---
title: <span style="color:red;">ANTs:</span><br /> A retrospective
author: Brian (Penn) and Nick (UVa)
bibliography: REFERENCES.bib
logo:    figures/antsLogoBase.png
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  structurebold
    highlight: tango
    incremental: yes
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
    toc: yes
  pdf_document:
    toc: true
    highlight: zenburn
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---
